---
title: "CorrelAid X Challenge: Import and Clean Unfallatlas Data"
author: "Cédric Scherer"
date: "`r Sys.Date()`"    ## current date
output:
  html_document:
    theme: paper          ## choose theme
    highlight: kate       ## choose coding style
    toc_depth: 2          ## table of content
    toc_float: true       ## allow toc next to doc#
    code_folding: show    ## include + show code
    #code_folding: hide    ## include code but hide it
    code_download: true   ## allow download of source
editor_options:
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = TRUE, warning = FALSE, 
  fig.showtext = TRUE, fig.retina = 1,
  fig.width = 10, fig.height = 7.5
)
```

```{r r-prep, message=FALSE}
library(tidyverse)
#library(readr)
library(here)
library(glue)
library(colorspace)
library(ragg)

theme_set(theme_minimal(base_size = 15, base_family = "Oswald"))

theme_update(
  axis.text = element_text(size = 11),
  axis.title.x = element_text(margin = margin(t = 10)),
  axis.title.y = element_text(margin = margin(r = 10)),
  legend.position = "top",
  panel.grid.major = element_line(size = .3),
  panel.grid.minor = element_blank()
)
```

[Data from Unfallatlas](https://unfallatlas.statistikportal.de/_opendata2020.html) comes as either tables (`.txt`) or shapefiles (`.shp`) 
for the years 2016 to 2019 for all Germany except Mecklenburg-Vorpommern (at least what we expect from the [interactive web map](https://unfallatlas.statistikportal.de/))


## IMPORT DATA

```{r}
## links to .csv .zip files
links <- c(
  "https://www.opengeodata.nrw.de/produkte/transport_verkehr/unfallatlas/Unfallorte2016_EPSG25832_CSV.zip",
  "https://www.opengeodata.nrw.de/produkte/transport_verkehr/unfallatlas/Unfallorte2017_EPSG25832_CSV.zip",
  "https://www.opengeodata.nrw.de/produkte/transport_verkehr/unfallatlas/Unfallorte2018_EPSG25832_CSV.zip",
  "https://unfallatlas.statistikportal.de/app/UnfalldatenDownload/Unfallorte2019_LinRef_csv.zip"
)

## .txt files of interest within each zip folder
files <- c(
  "csv/Unfallorte_2016_LinRef.txt",
  "csv/Unfallorte2017_LinRef.txt",
  "csv/Unfallorte2018_LinRef.txt",
  "csv/Unfallorte2019_LinRef.txt"
)

## function to download .zip to temp dir, extract + import data
read_multiple_zips <- function(link, file) {
    temp <- tempfile()
    download.file(link, temp)
    dat <- read_csv2(unz(temp, file))
    unlink(temp)
    dat
}

## process all 4 files + combine them in one df
df <- 
  map2_dfr(
    .x = links, 
    .y = files,
    .f = read_multiple_zips
  ) %>% 
  mutate_at(c(1, 8, 9), as.numeric) %>% 
  mutate(
    OBJECTID = if_else(is.na(OBJECTID), OBJECTID_1, OBJECTID),
    ULICHTVERH = if_else(is.na(ULICHTVERH), LICHT, ULICHTVERH),
    IstSonstige = if_else(is.na(IstSonstige), IstSonstig, IstSonstige)
  ) %>% 
  dplyr::select(-FID, -OBJECTID_1, -LICHT, -IstSonstig, -IstStrasse)
```

For the years 2016 to 2019, there are {nrow(df)} observations of all types of accidents with persons being injured ("Unfall mit Personenschaden").

## TRANSLATION

```{r}
## trnslation of column headers
## metadata: https://www.opengeodata.nrw.de/produkte/transport_verkehr/unfallatlas/DSB_Unfallatlas.pdf

df_eng <- 
  df %>% 
  rename(
    id = OBJECTID, 
    id_unknown = UIDENTSTLA, 
    federal_state = ULAND, 
    admin_district = UREGBEZ, 
    district = UKREIS, 
    municipality = UGEMEINDE, 
    year = UJAHR, 
    month = UMONAT, 
    hour = USTUNDE, 
    wday = UWOCHENTAG, ## 1: Sun; 2: Mon; ...
    category = UKATEGORIE, ## 1: dead; 2: serious injuries; 3: light injuries
    manner = UART, ## see metadata
    type = UTYP1,  ## see metadata
    is_bike = IstRad, 
    is_car = IstPKW, 
    is_pedestrian = IstFuss, 
    is_mtrcycle = IstKrad, 
    is_freightvehicle = IstGkfz,
    is_other = IstSonstige, 
    light_condition = ULICHTVERH, ## 0: day; 1: dawn/dusk; 2: dark
    street_condition = STRZUSTAND,  ## 0: dry; 1: wet; 2: icy
    x_lin = LINREFX, 
    y_lin = LINREFY, 
    x_wgs = XGCSWGS84, 
    y_wgs = YGCSWGS84
  ) %>% 
  dplyr::select(id:is_freightvehicle, is_other, street_condition, x_lin:y_wgs)

rds <- here("data", "unfallatlas_eng.Rds")
if(!file.exists(rds)) { write_rds(df_eng, rds)}
```


## BERLIN BIKE ACCIDENT DATA

```{r}
## Berlin and bike accidents only
df_bln_bikes <-
  df_eng %>% 
  filter(
    federal_state == "11",  ## Berlin only (13,390)
    is_bike == 1  ## Accidents involving bikes only (5,005)
  ) %>% 
  mutate(
    year = factor(year),
    month_lab = lubridate::month(month, label = T, abbr = F),
    wday_lab = lubridate::wday(wday, label = T, abbr = F)
  )

rds <- here("data", "unfallatlas_eng_berlin_bikes.Rds")
if(!file.exists(rds)) { write_rds(df_bln_bikes, rds) }
```


## DATA EXPLORATION

```{r}
df_bln_bikes %>% 
  group_by(year) %>% 
  count()
```


## PLOTS

## Accidents per year

```{r accidents-per-year, fig.width = 6, fig.height = 7}
df_bln_bikes %>% 
  count(year) %>% 
  ggplot(aes(year, n, fill = year)) +
    geom_col(width = .85) +
    geom_text(
      aes(label = n),
      family = "Oswald",
      fontface = "bold",
      color = "white",
      size = 4.5, 
      nudge_y = -200
    ) +
    scale_x_discrete(expand = c(.25, .25)) +
    scale_y_continuous(expand = c(0, 0)) +
    ggsci::scale_fill_jama(guide = F) +
    labs(x = NULL, y = "Anzahl der gemeldeten Unfälle") +
    theme(
      panel.grid.major.x = element_blank(),
      axis.text.x = element_text(size = 14, color = "black")
    ) +
    ggsave(here("plots", "accidents_year_bars.pdf"), width = 6, height = 7, device = cairo_pdf)

df_bln_bikes %>% 
  mutate(wend = ifelse(wday %in% c(1, 7), T, F)) %>% 
  count(year, wend) %>% 
  mutate(rel = if_else(wend == T, n / 2, n / 5)) %>% 
  ggplot(aes(wend, rel, fill = year)) +
    geom_col(position = position_dodge(width = .9), width = .85) +
    scale_x_discrete(labels = c("Werktags", "Wochenende"), expand = c(.25, .25)) +
    scale_y_continuous(expand = c(0, 0)) +
    ggsci::scale_fill_jama(name = "") +
    labs(x = NULL, y = "Anzahl der gemeldeten Unfälle pro Tag") +
    theme(
      panel.grid.major.x = element_blank(),
      axis.text.x = element_text(size = 14, color = "black")
    ) +
    ggsave(here("plots", "accidents_wday_year_bars_rel.pdf"), width = 8, height = 7, device = cairo_pdf)
```


## Accidents per year and month

```{r accidents-per-month-and-year, fig.width = 10, fig.height = 7}
df_bln_bikes %>% 
  count(year, month_lab) %>% 
  ggplot(aes(month_lab, n, fill = year)) +
    geom_col(position = "dodge", width = .8) +
    scale_y_continuous(
      expand = c(0, 0),
      breaks = 0:6 * 100
    ) +
    ggsci::scale_fill_jama(name = "") +
    labs(x = "Monat", y = "Anzahl der gemeldeten Unfälle") +
    theme(
      legend.position = c(.1, .81),
      panel.grid.major.x = element_blank()
    ) +
    ggsave(here("plots", "accidents_month_year_bars.pdf"), width = 10, height = 7, device = cairo_pdf)

df_bln_bikes %>% 
  count(year, month_lab) %>% 
  group_by(month_lab) %>% 
  mutate(top = if_else(n < max(n), 1.8, -.8)) %>% 
  ggplot(aes(month_lab, n, color = year, group = year)) +
    geom_line(
      aes(color = year, color = after_scale(colorspace::lighten(color, .55, space = "HLS"))),
      size = .8
    ) +
    geom_point(size = 2.3) +
    geom_text(
      aes(label = n, vjust = top),
      family = "Oswald",
      fontface = "bold",
      show.legend = F
    ) +
    scale_y_continuous(
      expand = c(0, 0),
      limits = c(0, 750),
      breaks = 0:6 * 100
    ) +
    ggsci::scale_color_jama(name = "") +
    labs(x = "Monat", y = "Anzahl der gemeldeten Unfälle") +
    theme(
      legend.position = c(.1, .78),
      panel.grid.major.x = element_blank()
    ) +
    ggsave(here("plots", "accidents_month_year_line.pdf"), width = 10, height = 7, device = cairo_pdf)
```

```{r accidents-per-month, fig.width = 10, fig.height = 7}
df_bln_bikes %>% 
  count(month_lab) %>% 
  ggplot(aes(month_lab, n, group = 1)) +
    geom_line(color = "grey85", size = .8) +
    geom_point(color = "grey30", size = 2.3) +
    geom_text(
      aes(label = n),
      family = "Oswald",
      fontface = "bold",
      color = "grey30",
      vjust = -.8, 
      show.legend = F
    ) +
    scale_y_continuous(
      expand = c(0, 0),
      limits = c(0, 1450),
      breaks = seq(0, 1200, by = 200)
    ) +
    labs(x = "Monat", y = "Anzahl der gemeldeten Unfälle") +
    theme(
      legend.position = c(.1, .78),
      panel.grid.major.x = element_blank()
    ) +
    ggsave(here("plots", "accidents_month_line.pdf"), width = 10, height = 7, device = cairo_pdf)
```

## Accidents per weekday

```{r accidents-per-wday-and-year, fig.width = 8, fig.height = 7}
df_bln_bikes %>% 
  mutate(wend = ifelse(wday %in% c(1, 7), T, F)) %>% 
  count(year, wend) %>% 
  ggplot(aes(wend, n, fill = year)) +
    geom_col(position = position_dodge(width = .9), width = .85) +
    scale_x_discrete(labels = c("Werktags", "Wochenende"), expand = c(.25, .25)) +
    scale_y_continuous(expand = c(0, 0)) +
    ggsci::scale_fill_jama(name = "") +
    labs(x = NULL, y = "Anzahl der gemeldeten Unfälle") +
    theme(
      panel.grid.major.x = element_blank(),
      axis.text.x = element_text(size = 14, color = "black")
    ) +
    ggsave(here("plots", "accidents_wday_year_bars.pdf"), width = 8, height = 7, device = cairo_pdf)

df_bln_bikes %>% 
  mutate(wend = ifelse(wday %in% c(1, 7), T, F)) %>% 
  count(year, wend) %>% 
  mutate(rel = if_else(wend == T, n / 2, n / 5)) %>% 
  ggplot(aes(wend, rel, fill = year)) +
    geom_col(position = position_dodge(width = .9), width = .85) +
    scale_x_discrete(labels = c("Werktags", "Wochenende"), expand = c(.25, .25)) +
    scale_y_continuous(expand = c(0, 0)) +
    ggsci::scale_fill_jama(name = "") +
    labs(x = NULL, y = "Anzahl der gemeldeten Unfälle pro Tag") +
    theme(
      panel.grid.major.x = element_blank(),
      axis.text.x = element_text(size = 14, color = "black")
    ) +
    ggsave(here("plots", "accidents_wday_year_bars_rel.pdf"), width = 8, height = 7, device = cairo_pdf)
```


### Accidents per hour

```{r accidents-per-hour-and-year, fig.width = 10, fig.height = 7}
df_bln_bikes %>% 
  mutate(hour = factor(hour)) %>% 
  count(year, hour) %>% 
  ggplot(aes(hour, n, fill = year)) +
    geom_col(position = "dodge", width = .8) +
    scale_y_continuous(expand = c(0, 0)) +
    ggsci::scale_fill_jama(name = "") +
    labs(x = "Uhrzeit", y = "Anzahl der gemeldeten Unfälle") +
    theme(
      legend.position = c(.1, .81),
      panel.grid.major.x = element_blank()
    ) +
    ggsave(here("plots", "accidents_hour_year_bars.pdf"), width = 10, height = 7, device = cairo_pdf)

df_bln_bikes %>% 
  mutate(hour = factor(hour)) %>% 
  count(year, hour) %>% 
  group_by(hour) %>% 
  mutate(top = if_else(n < max(n), 1.8, -.8)) %>% 
  ggplot(aes(hour, n, color = year, group = year)) +
    geom_line(
      aes(color = year, color = after_scale(colorspace::lighten(color, .55, space = "HLS"))),
      size = .8
    ) +
    geom_point(size = 2.3) +
    geom_text(
      aes(label = n, vjust = top),
      family = "Oswald",
      fontface = "bold",
      size = 2.5,
      show.legend = F
    ) +
    coord_cartesian(clip = "off") +
    scale_y_continuous(
      expand = c(.002, .002),
      limits = c(0, 550)
    ) +
    ggsci::scale_color_jama(name = "") +
    labs(x = "Uhrzeit", y = "Anzahl der gemeldeten Unfälle") +
    theme(
      legend.position = c(.1, .78),
      panel.grid.major.x = element_blank()
    ) +
    ggsave(here("plots", "accidents_hour_year_line.pdf"), width = 10, height = 7, device = cairo_pdf)
```

```{r accidents-per-hour, fig.width = 10, fig.height = 7}
df_bln_bikes %>% 
  mutate(hour = factor(hour)) %>% 
  count(hour) %>% 
  ggplot(aes(hour, n)) +
    geom_col(width = .8, color = "grey30") +
    scale_y_continuous(expand = c(0, 0)) +
    ggsci::scale_fill_jama(name = "") +
    labs(x = "Uhrzeit", y = "Anzahl der gemeldeten Unfälle") +
    theme(panel.grid.major.x = element_blank()) +
    ggsave(here("plots", "accidents_hour_bars.pdf"), width = 10, height = 7, device = cairo_pdf)

df_bln_bikes %>% 
  mutate(hour = factor(hour)) %>% 
  count(hour) %>% 
  ggplot(aes(hour, n, group = 1)) +
    geom_line(color = "grey85", size = .8) +
    geom_point(color = "grey30", size = 2.3) +
    geom_text(
      aes(label = n),
      family = "Oswald",
      fontface = "bold",
      color = "grey30",
      size = 2.5,
      vjust = -.8,
      show.legend = F
    ) +
    coord_cartesian(clip = "off") +
    scale_y_continuous(
      expand = c(.002, .002),
      limits = c(0, 1060),
      breaks = 0:5 * 200
    ) +
    ggsci::scale_color_jama(name = "") +
    labs(x = "Uhrzeit", y = "Anzahl der gemeldeten Unfälle") +
    theme(
      legend.position = c(.1, .78),
      panel.grid.major.x = element_blank()
    ) +
    ggsave(here("plots", "accidents_hour_line.pdf"), width = 10, height = 7, device = cairo_pdf)
```

### Heatmap

```{r}
df_bln_bikes %>% 
  count(wday_lab, month_lab, year) %>% 
  ggplot(aes(wday_lab, fct_rev(month_lab))) +
    geom_tile(aes(fill = n), color = "white", size = .4) +
    facet_wrap(~year) +
    scale_x_discrete(expand = c(0, 0)) +
    scale_y_discrete(expand = c(0, 0)) +
    scico::scale_fill_scico(
      palette = "bilbao", 
      name = "Gemeldete Unfälle",
      limits = c(0, NA)
    ) +
    guides(fill = guide_colourbar(title.position = "top",
                                  title.hjust = .5,
                                  label.hjust = .5,
                                  barwidth = unit(20, "lines"),
                                  barheight = unit(.4, "lines"))) +
    labs(x = NULL, y = NULL) +
    theme(
      panel.grid.major = element_blank(),
      panel.background = element_rect(color = "grey30", fill = "transparent", size = .8),
      panel.ontop = T,
      panel.spacing = unit(1.3, "lines"),
      strip.text = element_text(size = 15, color = "black", face = "bold"),
      legend.title = element_text(size = 12, color = "grey30"),
      legend.text = element_text(size = 10, color = "grey30"),
      legend.position = "bottom"
    ) +
    ggsave(here("plots", "accidents_heatmap.pdf"), width = 10, height = 7, device = cairo_pdf)
```



```{r convert-to-png}
## convert PDFs to PNGs
pdfs <- list.files(here("plots"), pattern = ".*pdf")
setwd(here("plots", "png"))
for(pdf in pdfs) {
  pdftools::pdf_convert(pdf = glue("{here('plots')}/{pdf}"), 
                        format = "png", dpi = 750)
}
```
